Escaping RGBland: Selecting Colors for Statistical Graphics

### 2008-11-19 pkg/paper/Palettes/CSDA.Rnw
Escaping RGBland paper accepted for CSDA. Small touches to the paper, changed references in the package. Prepared release 1.0-0 accompanying the CSDA paper.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@50 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[shortnames]{Z}

@@

\Abstract{
  Statistical graphics are often augmented by the use of color
  coding information contained in some variable. When this involves
  the shading of areas (and not only points or lines)---e.g., as in bar plots, pie charts,
  mosaic displays or heatmaps---it is important that the colors are perceptually based
  and do not introduce optical illusions or systematic bias. This article discusses
  how the perceptually-based Hue-Chroma-Luminance (HCL) color space can be used for deriving suitable
  color palettes for coding categorical data (qualitative palettes) and numerical
  variables (sequential and diverging palettes).
}

@@

The \proglang{R} system for statistical computing \citep{color:R:2008} provides
an open-source implementation of HCL (and other color spaces) in the package
\pkg{colorspace}, originally written by Ross Ihaka. The coordinate transformations mentioned
above are contained in \proglang{C}~code within \pkg{colorspace} that are easy to
port to other statistical software systems. Starting from version~0.97, %% FIXME
\pkg{colorspace} \citep{color:Ihaka+Murrell+Hornik+Zeileis:2008} also includes
an implementation of all palettes discussed above. 
\citep[Originally, the code for the palettes was in the \pkg{vcd} package,]%
[but it was recently moved to \pkg{colorspace} to be more easily acessible.]%
{color:Meyer+Zeileis+Hornik:2008} Qualitative palettes are provided by
\code{rainbow_hcl()} (named after the HSV-based function \code{rainbow()} in base
\proglang{R}). Sequential palettes based on a single hue are implemented in the function
\code{sequential_hcl()} while \code{heat_hcl()} offers sequential palettes based on
a range of hues. Diverging palettes can be obtained by \code{diverge_hcl()}.
Technical documentation along with a large collection of example palettes is available via
\verb/help("rainbow_hcl", package = "colorspace")/. Furthermore, \proglang{R}~code for reproducing
Figures~\ref{fig:pal-q-examples}, \ref{fig:pal-s-examples} and \ref{fig:pal-d-examples}
(and some illustrations) can be accessed via \code{vignette("hcl-colors", package = "colorspace")}.

>>>>

\documentclass[shortnames,noheadings]{Z}

@@

\Abstract{
  Statistical graphics are often augmented by the use of color
  coding information contained in some variable. When this involves
  the shading of areas (and not only points or lines)---e.g., as in bar plots, pie charts,
  mosaic displays or heatmaps---it is important that the colors are perceptually based
  and do not introduce optical illusions or systematic bias.
  Based on the perceptually-based Hue-Chroma-Luminance (HCL) color space  
  suitable color palettes are derived for coding categorical data (qualitative palettes) and numerical
  variables (sequential and diverging palettes).
}

\usepackage{fancyhdr}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0pt}
\pagestyle{fancy}
\thispagestyle{plain}
\fancyhf{}
\fancyhead[RO,LE]{\thepage} 
\fancyhead[CO]{\textit{\normalsize Achim Zeileis, Kurt Hornik, Paul Murrell}} 
\fancyhead[CE]{\textit{\normalsize Escaping RGBland: Selecting Colors for Statistical Graphics}} 
\fancyfoot[LO,LE]{\small Copyright {\copyright} 2008 Elsevier B.V.}
\fancypagestyle{plain}{
\fancyhf{}
\fancyfoot[LO,LE]{\small This is a preprint of an article accepted for publication in
\textit{Computational Statistics \& Data Analysis}.\\
 Copyright {\copyright} 2008 Elsevier B.V.} 
}

@@

The \proglang{R} system for statistical computing \citep{color:R:2008} provides
an open-source implementation of HCL (and other color spaces) in the package
\pkg{colorspace}, originally written by Ross Ihaka. The coordinate transformations mentioned
above are contained in \proglang{C}~code within \pkg{colorspace} that are easy to
port to other statistical software systems. Version~1.0-0 of
\pkg{colorspace} \citep{color:Ihaka+Murrell+Hornik+Zeileis:2008} also includes
an implementation of all palettes discussed above. 
\citep[Originally, the code for the palettes was in the \pkg{vcd} package,]%
[but it was recently moved to \pkg{colorspace} to be more easily acessible.]%
{color:Meyer+Zeileis+Hornik:2006} Qualitative palettes are provided by
\code{rainbow_hcl()} (named after the HSV-based function \code{rainbow()} in base
\proglang{R}). Sequential palettes based on a single hue are implemented in the function
\code{sequential_hcl()} while \code{heat_hcl()} offers sequential palettes based on
a range of hues. Diverging palettes can be obtained by \code{diverge_hcl()}.
Technical documentation along with a large collection of example palettes is available via
\verb/help("rainbow_hcl", package = "colorspace")/. Furthermore, \proglang{R}~code for reproducing
Figures~\ref{fig:pal-q-examples}, \ref{fig:pal-s-examples} and \ref{fig:pal-d-examples}
(and some illustrations) can be accessed via \code{vignette("hcl-colors", package = "colorspace")}.
